---
title: "Ma_Lab_RNA_Seq_Bulk_Beitong_EM_1M1"
output: html_document
date: "2023-08-29"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
setwd("/Users/ankitverma/Documents/rna-seq/ma_lab_dm6/RNA_EM_1M1")

```{r library, message=FALSE}
library(DESeq2)
library(ggplot2)
library(pheatmap)
library(plyr)
library(ggpubr)
library(splitstackshape)
library(gprofiler2)
library(UpSetR)
library(tidyr)
library(org.Dm.eg.db)
library(stringr)
library(ggrepel)
library(ChIPseeker)
library(TxDb.Dmelanogaster.UCSC.dm6.ensGene)
library(clusterProfiler)
library(stringi)
library(data.table)
library(openxlsx)
```


### Obtain un-normalized counts from beitong_em_dds (output of nextflow nf-core/rnaseq)
```{r Obtaining un-normalized counts from dds}
load("/Users/ankitverma/Documents/rna-seq/ma_lab_dm6/RNA_EM_1M1/beitong_deseq2.dds.RData")
beitong_em_dds <- dds
rm(dds)
beitong_em_precountdata <- counts(beitong_em_dds, normalized=FALSE)
beitong_em_precountdata <- data.frame(beitong_em_precountdata)
beitong_em_precountdata["ensID"] <- rownames(beitong_em_precountdata)
write.table(beitong_em_precountdata, "beitong_em_precountdata.txt", quote = F, append = F, sep="\t")
colnames(beitong_em_precountdata)[1] <- c("EM_1M1")
```

### Obtain un-normalized counts from ason_all_dds (output of nextflow nf-core/rnaseq)
```{r Obtaining un-normalized counts from ason dds}
load("/Users/ankitverma/Documents/rna-seq/ma_lab_dm6/RNA_EM_1M1/ason_and_all_deseq2.dds.RData")
ason_all_dds <- dds
rm(dds)
ason_all_precountdata <- counts(ason_all_dds, normalized=FALSE)
ason_all_precountdata <- data.frame(ason_all_precountdata)
ason_all_precountdata["ensID"] <- rownames(ason_all_precountdata)
write.table(ason_all_precountdata, "ason_all_precountdata.txt", quote = F, append = F, sep="\t")
```


```{r Merge ason and beitong data}
beitong_ason_precountdata <- merge(beitong_em_precountdata, ason_all_precountdata, by="ensID")
```

# Merge old(run1) + new(run2) Beitong counts + new (run3) Beitong and Ason (run1)
```{r Merge old and new data counts}
precountdata_beitong_old1_new2 <- read.table("/Users/ankitverma/Documents/rna-seq/ma_lab_dm6/RNAseq_with_Adam/precountdata_beitong_old_new.txt")

# remove run1 samples as they are already being imported

precountdata_beitong_new2 <- precountdata_beitong_old1_new2[,c(1:4)]
precountdata_beitong_old1_new2_new3_ason1 <- merge(precountdata_beitong_new2, beitong_ason_precountdata, by="ensID")

#renmae it to smaller id 
pbonna <- precountdata_beitong_old1_new2_new3_ason1
```


### Prepare new count data for beitong merged data
```{r Preparing count matrix beitong}
rownames(pbonna) <- pbonna$ensID
pbonna_countdata = as.matrix(pbonna[,c(2:9,12:13)])
pbonna_coldata <- data.frame(colnames(pbonna_countdata))
colnames(pbonna_coldata) <- c("sample")

write.table(pbonna_countdata, "pbonna_countdata.txt", quote = F, append = F, sep="\t")
```


### Prepare new count matrix for beitong merged data
```{r Preparing count matrix nuria}
all(rownames(pbonna_coldata) == colnames(pbonna_countdata)) #should print TRUE
beitong_ddsO <- DESeqDataSetFromMatrix(countData = pbonna_countdata, colData = pbonna_coldata, design = ~ Sample_Status)
beitong_keep <- rowSums(counts(beitong_ddsO)) > 10
beitong_ddsO <- beitong_ddsO[beitong_keep,]
